---
title: "Frequency Analysis of the Injury Reports"
author: "Abdullah Alsharef"
date: "3/20/2020"
header-includes:
- \usepackage{afterpage}
output:
  pdf_document:
    latex_engine: xelatex
    df_print: kable
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    number_sections: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r, echo=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(janitor)
library(maps)
library(knitr)
```


\newpage

# Introduction
The goal of the this document is to prepare the Injury reports for ML/NLP analysis.  This document shows the frequency of the targeted variables or attributes.

Figure 1 shows an overview of suggested process.

We are attempting to test the concept of predicting the attributes of injury reports from the injury's narratives (The textual unstructured description).  This is a supervised learning problem.  The attributes of the injury report are:

- Event
- Source
- Amputation
- Body Part
- Contractor
- Hospitalization
- Nature

The injury narrative's is the predictor (x variable) and the attributes (listed above) are the (Y variable).  

In this folder, there are separate sub folders with the name of the attributes.  Each sub folder has a CSV file.  Each CSV have 3 columns (ID Number - Narratives - The labeled attribute).  The number of reports in each folder is different because we removed all unlabeled data.

```{r label, out.width = "75%", fig.cap = "NLP/ML Overview",fig.align='center'} 
knitr::include_graphics("OSHA Severe ML_NLP - Page 2.pdf")

```
\newpage


# Frequency Analysis 


```{r}
load("E:/Research/Safety Research 2020/Severe Injury_OSHADATA/Datasets/R Project/OSHA_Severe/Final_Prepetation/cons_v2_.RData")

# filter the needed variables

cons_v3 <- cons_v2 %>% 
  select(ID,Final_Narrative, NAICS_L3, NAICS_L4, NAICS_L5,	event_L1,	event_L2,	event_L3,	part_L2,	part_L3,	part_L1,	nature_L1,	nature_L2,	nature_L3,	source1_L1,	source1_L2,	source1_L3,	source2_L1,	source2_L2,	source2_L3,	Hospitalized_Y_N,	Amputation_Y_N)

```

## Event

The number of reports (after excludeing unlabeled reports) is **8,472** reports. The number of categories for *Event* is **7**.

```{r}

cons_event <- cons_v3 %>% 
  select(ID, Final_Narrative, event_L1) %>% 
  mutate(Event = tolower(event_L1) ) %>% 
  filter(Event != "nonclassifiable") 

cons_event_count <- cons_event %>% 
  count(Event, sort = T) %>% 
  mutate(Event_p = ((n*100)/sum(n)))

cons_event_count


# write.csv(cons_event, "Final_Prepetation/cons_event_1.csv", na = "")
```

## Source

The number of reports (after excludeing unlabeled reports) is **8,143** reports. The number of categories for Source is **9**.

```{r}
cons_Source <- cons_v3 %>% 
  select(ID, Final_Narrative, source1_L1) %>% 
  mutate(Source = tolower(source1_L1) ) %>% 
  filter(Source != "nonclassifiable") 

cons_Source %>% 
  count(Source, sort = T) %>% 
   mutate(Source_p = ((n*100)/sum(n)))

# write.csv(cons_Source, "Final_Prepetation/cons_Source.csv", na = "")

```

## Amputation

The number of reports (after excludeing unlabeled reports) is **8,563 ** reports.  This is similar to sentiment analysis in which the Y variable (the response) is either "Yes" or "No".

```{r}
amp <- cons_v3 %>% 
  select(ID,Final_Narrative, Amputation_Y_N)

amp %>% 
  count(Amputation_Y_N, sort = T) %>% 
   mutate(Amputation_p = ((n*100)/sum(n)))

# write.csv(hosp, "Final_Prepetation/amp.csv", na = "")

```

## Body Part
The number of reports (after excludeing unlabeled reports) is **7,897** reports. The number of categories for Source is **6**.

```{r}
# We'll use level 1. We'll merge "Head" with "Neck". 
cons_part <- cons_v3 %>% 
  select(ID, Final_Narrative, part_L1) %>% 
  mutate(Part = tolower(part_L1) ) %>% 
  filter(Part != "nonclassifiable") %>% 
  mutate(Part_v2 = case_when(
    Part == "head" ~ "head and neck",
    Part == "neck, including throat" ~ "head and neck",
    TRUE ~ as.character(Part))) %>% 
  select(-Part,-part_L1)


cons_part %>% 
  count(Part_v2, sort = T) %>% 
   mutate(Part_v2_p = ((n*100)/sum(n)))

# write.csv(cons_part, "Final_Prepetation/cons_part.csv", na = "")

```

## Contractor
The number of reports (after excludeing unlabeled reports) is **8,561** reports. The number of categories for Source is **9**.

```{r}
# the contractor variable
#We'll use level 2 in prediction as an attribute or predictor.

# - We'll use level 2.
# - NAICS_L4  == "Construction" ~ NA_character_ ,
# - NAICS_L4  == "Specialty Trade Contractors" ~ "Other Specialty Trade Contractors",
# - NAICS_L4  == "Land Subdivision" ~ "Other Heavy and Civil Engineering Construction"
# 
# Discard the two construction cases and do the prediction.
# 
# 

cons_cont <- cons_v3 %>% 
  select(ID, Final_Narrative, NAICS_L4) %>% 
  mutate( Contractor = case_when(
   NAICS_L4  == "Construction" ~ NA_character_ ,
   NAICS_L4  == "Specialty Trade Contractors" ~ "Other Specialty Trade Contractors",
   NAICS_L4  == "Land Subdivision" ~ "Other Heavy and Civil Engineering Construction",
    TRUE ~ as.character(NAICS_L4)))  %>% 
  filter(!is.na(Contractor)) %>% 
  select(-NAICS_L4)

cons_cont %>% 
  count(Contractor, sort = T) %>% 
   mutate(Contractor_p = ((n*100)/sum(n)))

# cons_cont %>% 
#   count(Contractor) %>% 
#   mutate(len = length(unique(Contractor)))

# Output the data set

#write.csv(cons_cont, "Final_Prepetation/Contractor.csv", na = "")


```

```{r}
##V2

# - We'll use level 2 with combining some categories together.
# - We'll call general categories "Other ...."
# - Exclude non-classifiable

cons_event_v2 <- cons_v3 %>% 
  select(ID, Final_Narrative, event_L2) %>% 
  mutate( Event_v2 = case_when(
   event_L2 == "Fall, slip, trip, unspecified" ~ "Other - falls, slips, trips",
   event_L2 == "Jumps to lower level" ~ "Other - falls, slips, trips",
   event_L2 == "Jumps to lower level" ~ "Other - falls, slips, trips",
   event_L2 == "Slip or trip without fall" ~ "Other - falls, slips, trips",
   event_L2 == "Fall or jump curtailed by personal fall arrest system" ~ "Other - falls, slips, trips",
   event_L2 == "Fall, slip, trip, n.e.c." ~ "Other - falls, slips, trips",
   event_L2 == "Contact with objects and equipment, unspecified" ~ "Other - contact with objects and equipment",
   event_L2 == "Struck, caught, or crushed in collapsing structure, equipment, or material" ~ "Other - contact with objects and equipment",
   event_L2 == "Contact with objects and equipment, n.e.c." ~ "Other - contact with objects and equipment",
   event_L2 == "Rubbed or abraded by friction or pressure" ~ "Other - contact with objects and equipment",
   event_L2 == "Pedestrian vehicular incidents" ~ "transportation incidents",
   event_L2 == "Nonroadway incidents involving motorized land vehicles" ~ "transportation incidents",
   event_L2 == "Roadway incidents involving motorized land vehicle" ~ "transportation incidents",
   event_L2 == "Water vehicle incidents" ~ "transportation incidents",
   event_L2 == "Transportation incident, unspecified" ~ "transportation incidents",
   event_L2 == "Animal and other non-motorized vehicle transportation incidents" ~ "transportation incidents",
   event_L2 == "Aircraft incidents" ~ "transportation incidents",
   event_L2 == "Rail vehicle incidents" ~ "transportation incidents",
   event_L2 == "Fires" ~ "fires and explosions",
  event_L2 == "Explosions" ~ "fires and explosions",
  event_L2 == "Fire or explosion, unspecified" ~ "fires and explosions",
  event_L2 == "Overexertion involving outside sources" ~ "overexertion and bodily reaction",
  event_L2 == "Other exertions or bodily reactions" ~ "overexertion and bodily reaction",
  event_L2 == "Overexertion and bodily reaction, unspecified" ~ "overexertion and bodily reaction",
  event_L2 == "Bodily conditions, n.e.c." ~ "overexertion and bodily reaction",
  event_L2 == "Animal and insect related incidents" ~ "violence and other injuries by persons or animals",
  event_L2 == "Intentional injury by person" ~ "violence and other injuries by persons or animals",
  event_L2 == "Injury by person—unintentional or intent unknown" ~ "violence and other injuries by persons or animals",
  event_L2 == "Exposure to other harmful substances" ~ "exposure to harmful substances or environments",
  event_L2 == "Exposure to harmful substances or environments, unspecified" ~ "exposure to harmful substances or environments",
  TRUE ~ as.character(event_L2)))  %>% 
  filter(event_L2 != "Nonclassifiable")


cons_event_v2_count <- cons_event_v2 %>% 
  count(Event_v2, sort = T)


# write.csv(cons_event_v2, "Final_Prepetation/cons_event_v2.csv", na = "")



```


## Hospitalization
The number of reports (after excludeing unlabeled reports) is **8,563** reports.  This is similar to sentiment analysis in which the Y variable is either "Yes" or "No".

```{r}
hosp <- cons_v3 %>% 
  select(ID,Final_Narrative, Hospitalized_Y_N)


hosp %>% 
  count(Hospitalized_Y_N, sort = T) %>% 
   mutate(Hospitalized_p = ((n*100)/sum(n)))

#write.csv(hosp, "Final_Prepetation/hosp.csv", na = "")

```

## Nature

The number of reports (after excludeing unlabeled reports) is **8,535** reports. The number of categories for Source is **8**.


```{r}
# we'll use level 2 since the majority of cases occurred for TRAUMATIC INJURIES AND DISORDERS with 8535 cases.

# Note: # since the count is small, we'll remove "Other - Nature" 

cons_nature <- cons_v3 %>% 
  select(ID, Final_Narrative, nature_L2) %>% 
  mutate( Nature = case_when(
   nature_L2 == "Surface wounds and bruises" ~ "Open wounds",
   nature_L2 == "Traumatic injuries and disorders, unspecified" ~ "Other traumatic injuries and disorders",
    nature_L2 == "Circulatory system diseases" ~ "Other - Nature",
   nature_L2 == "Respiratory system diseases" ~ "Other - Nature",
   nature_L2 == "Disorders of the skin and subcutaneous tissue" ~ "Other - Nature",
   nature_L2 == "Symptoms" ~ "Other - Nature",
   nature_L2 == "EXPOSURES TO DISEASE—NO ILLNESS INCURRED" ~ "Other - Nature",
   nature_L2 == "Bacterial diseases" ~ "Other - Nature",
   TRUE ~ as.character(nature_L2)))

# cons_nature %>% 
#   count(Nature, sort = T)

# since the count is small, we'll remove "Other - Nature" 
cons_nature_2 <- cons_nature %>% 
  select(-nature_L2) %>% 
  filter (Nature!= "Other - Nature")

cons_nature_2 %>% 
  count(Nature, sort = T) %>% 
   mutate(Nature_p = ((n*100)/sum(n)))

# write.csv(cons_nature_2, "Final_Prepetation/cons_nature_2.csv", na = "")

```






















