---
title: "Preparing the Severe Injury Reports for ML"
author: "Abdullah Alsharef"
date: "3/20/2020"
output:
  pdf_document:
    df_print: kable
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    number_sections: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r, echo=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(janitor)
library(maps)
```


\newpage

# Introduction
The goal of the this document is to prepare the OSHa Severe Database for ML/NLP analysis. We are going to modify the categories to fit the criteria of the big dimenstionality.

# Contractor

We'll use level 2 in prediction as an attribute or predictor.

- We'll use level 2.
- NAICS_L4  == "Construction" ~ NA_character_ ,
- NAICS_L4  == "Specialty Trade Contractors" ~ "Other Specialty Trade Contractors",
- NAICS_L4  == "Land Subdivision" ~ "Other Heavy and Civil Engineering Construction"

Discard the two construction cases and do the prediction.

```{r}
load("E:/Research/Safety Research 2020/Severe Injury_OSHADATA/Datasets/R Project/OSHA_Severe/Final_Prepetation/cons_v2_.RData")

# filter the needed variables

cons_v3 <- cons_v2 %>% 
  select(ID,Final_Narrative, NAICS_L3, NAICS_L4, NAICS_L5,	event_L1,	event_L2,	event_L3,	part_L2,	part_L3,	part_L1,	nature_L1,	nature_L2,	nature_L3,	source1_L1,	source1_L2,	source1_L3,	source2_L1,	source2_L2,	source2_L3,	Hospitalized_Y_N,	Amputation_Y_N)

# the contractor variable
cons_cont <- cons_v3 %>% 
  select(ID, Final_Narrative, NAICS_L4) %>% 
  mutate( Contractor = case_when(
   NAICS_L4  == "Construction" ~ NA_character_ ,
   NAICS_L4  == "Specialty Trade Contractors" ~ "Other Specialty Trade Contractors",
   NAICS_L4  == "Land Subdivision" ~ "Other Heavy and Civil Engineering Construction",
    TRUE ~ as.character(NAICS_L4)))  %>% 
  filter(!is.na(Contractor)) %>% 
  select(-NAICS_L4)

cons_cont %>% 
  count(Contractor) %>% 
  mutate(len = length(unique(Contractor)))

# Output the data set

#write.csv(cons_cont, "Final_Prepetation/Contractor.csv", na = "")


```


# Event

We'll use Level 1.  Exclude all nonclassifiable cases - n = 91 cases.

## V1

```{r}

cons_event <- cons_v3 %>% 
  select(ID, Final_Narrative, event_L1) %>% 
  mutate(Event = tolower(event_L1) ) %>% 
  filter(Event != "nonclassifiable") 

cons_event_count <- cons_event %>% 
  count(Event, sort = T)


# write.csv(cons_event, "Final_Prepetation/cons_event_1.csv", na = "")
```


##V2

- We'll use level 2 with combining some categories together.
- We'll call general categories "Other ...."
- Exclude nonclassifiable

```{r}


cons_event_v2 <- cons_v3 %>% 
  select(ID, Final_Narrative, event_L2) %>% 
  mutate( Event_v2 = case_when(
   event_L2 == "Fall, slip, trip, unspecified" ~ "Other - falls, slips, trips",
   event_L2 == "Jumps to lower level" ~ "Other - falls, slips, trips",
   event_L2 == "Jumps to lower level" ~ "Other - falls, slips, trips",
   event_L2 == "Slip or trip without fall" ~ "Other - falls, slips, trips",
   event_L2 == "Fall or jump curtailed by personal fall arrest system" ~ "Other - falls, slips, trips",
   event_L2 == "Fall, slip, trip, n.e.c." ~ "Other - falls, slips, trips",
   event_L2 == "Contact with objects and equipment, unspecified" ~ "Other - contact with objects and equipment",
   event_L2 == "Struck, caught, or crushed in collapsing structure, equipment, or material" ~ "Other - contact with objects and equipment",
   event_L2 == "Contact with objects and equipment, n.e.c." ~ "Other - contact with objects and equipment",
   event_L2 == "Rubbed or abraded by friction or pressure" ~ "Other - contact with objects and equipment",
   event_L2 == "Pedestrian vehicular incidents" ~ "transportation incidents",
   event_L2 == "Nonroadway incidents involving motorized land vehicles" ~ "transportation incidents",
   event_L2 == "Roadway incidents involving motorized land vehicle" ~ "transportation incidents",
   event_L2 == "Water vehicle incidents" ~ "transportation incidents",
   event_L2 == "Transportation incident, unspecified" ~ "transportation incidents",
   event_L2 == "Animal and other non-motorized vehicle transportation incidents" ~ "transportation incidents",
   event_L2 == "Aircraft incidents" ~ "transportation incidents",
   event_L2 == "Rail vehicle incidents" ~ "transportation incidents",
   event_L2 == "Fires" ~ "fires and explosions",
  event_L2 == "Explosions" ~ "fires and explosions",
  event_L2 == "Fire or explosion, unspecified" ~ "fires and explosions",
  event_L2 == "Overexertion involving outside sources" ~ "overexertion and bodily reaction",
  event_L2 == "Other exertions or bodily reactions" ~ "overexertion and bodily reaction",
  event_L2 == "Overexertion and bodily reaction, unspecified" ~ "overexertion and bodily reaction",
  event_L2 == "Bodily conditions, n.e.c." ~ "overexertion and bodily reaction",
  event_L2 == "Animal and insect related incidents" ~ "violence and other injuries by persons or animals",
  event_L2 == "Intentional injury by person" ~ "violence and other injuries by persons or animals",
  event_L2 == "Injury by person—unintentional or intent unknown" ~ "violence and other injuries by persons or animals",
  event_L2 == "Exposure to other harmful substances" ~ "exposure to harmful substances or environments",
  event_L2 == "Exposure to harmful substances or environments, unspecified" ~ "exposure to harmful substances or environments",
  TRUE ~ as.character(event_L2)))  %>% 
  filter(event_L2 != "Nonclassifiable")


cons_event_v2_count <- cons_event_v2 %>% 
  count(Event_v2, sort = T)


# write.csv(cons_event_v2, "Final_Prepetation/cons_event_v2.csv", na = "")



```


# Hospitilization

```{r}
hosp <- cons_v3 %>% 
  select(ID,Final_Narrative, Hospitalized_Y_N)

#write.csv(hosp, "Final_Prepetation/hosp.csv", na = "")

```


# Amputation


```{r}
amp <- cons_v3 %>% 
  select(ID,Final_Narrative, Amputation_Y_N)

# write.csv(hosp, "Final_Prepetation/amp.csv", na = "")

```

# Source

We'll use level 1. Exlude "Nonclassifiable".

```{r}
cons_Source <- cons_v3 %>% 
  select(ID, Final_Narrative, source1_L1) %>% 
  mutate(Source = tolower(source1_L1) ) %>% 
  filter(Source != "nonclassifiable") 

cons_Source %>% 
  count(Source, sort = T)

# write.csv(cons_Source, "Final_Prepetation/cons_Source.csv", na = "")

```


# Body Part

We'll use level 1. We'll merge "Head" with "Neck". 

```{r}
cons_part <- cons_v3 %>% 
  select(ID, Final_Narrative, part_L1) %>% 
  mutate(Part = tolower(part_L1) ) %>% 
  filter(Part != "nonclassifiable") %>% 
  mutate(Part_v2 = case_when(
    Part == "head" ~ "head and neck",
    Part == "neck, including throat" ~ "head and neck",
    TRUE ~ as.character(Part))) %>% 
  select(-Part,-part_L1)

# write.csv(cons_part, "Final_Prepetation/cons_part.csv", na = "")

```


# Nature

we'll use level 2 since the majority of cases occured for TRAUMATIC INJURIES AND DISORDERS with 8535 cases.



```{r}
cons_nature <- cons_v3 %>% 
  select(ID, Final_Narrative, nature_L2) %>% 
  mutate( Nature = case_when(
   nature_L2 == "Surface wounds and bruises" ~ "Open wounds",
   nature_L2 == "Traumatic injuries and disorders, unspecified" ~ "Other traumatic injuries and disorders",
    nature_L2 == "Circulatory system diseases" ~ "Other - Nature",
   nature_L2 == "Respiratory system diseases" ~ "Other - Nature",
   nature_L2 == "Disorders of the skin and subcutaneous tissue" ~ "Other - Nature",
   nature_L2 == "Symptoms" ~ "Other - Nature",
   nature_L2 == "EXPOSURES TO DISEASE—NO ILLNESS INCURRED" ~ "Other - Nature",
   nature_L2 == "Bacterial diseases" ~ "Other - Nature",
   TRUE ~ as.character(nature_L2)))

cons_nature %>% 
  count(Nature, sort = T)

# since the count is small, we'll remove "Other - Nature" 
cons_nature_2 <- cons_nature %>% 
  select(-nature_L2) %>% 
  filter (Nature!= "Other - Nature")

# write.csv(cons_nature_2, "Final_Prepetation/cons_nature_2.csv", na = "")

```






















